---
title: "rsingav1_4"
output: word_document
date: "2023-11-20"
---



```{r}
# Import the lpSolveAPI library
library(lpSolveAPI)
```

# Decision variables:
Let A1 be the number of workers assigned to shift 1
Let A2 be the number of workers assigned to shift 2
Let A3 be the number of workers assigned to shift 3
Let A4 be the number of workers assigned to shift 4
Let A5 be the number of workers assigned to shift 5
Let A6 be the number of workers assigned to shift 6
Let A7 be the number of workers assigned to shift 7

# Objective Function:
Minimize the total wages:

775*A1 + 800*A2 + 800*A3 + 800*A4 + 800*A5 + 775*A6 + 750*A7

```{r}
# Create a new lp problem
AP_IP_model = make.lp(0, 7)

# Set the objective function coefficients(decision variables)
set.objfn(AP_IP_model, c(775, 800, 800, 800, 800, 775, 750))

```

# Subject to Constraints:
A2+A3+A4+A5+A6 >=18 for Sunday
A3+A4+A5+A6+A7 >=27 for Monday
A1+A4+A5+A6+A7 >=22 for Tuesday
A1+A2+A5+A6+A7 >=26 for Wednesday
A1+A2+A3+A6+A7 >=25 for Thursday
A1+A2+A3+A4+A7 >=21 for Friday
A1+A2+A3+A4+A5 >=19 for Saturday

```{r}
# Add constraints
add.constraint(AP_IP_model, c(0, 1, 1, 1, 1, 1, 0), ">=", 18)  # Sunday
add.constraint(AP_IP_model, c(0, 0, 1, 1, 1, 1, 1), ">=", 27)  # Monday
add.constraint(AP_IP_model, c(1, 0, 0, 1, 1, 1, 1), ">=", 22)  # Tuesday
add.constraint(AP_IP_model, c(1, 1, 0, 0, 1, 1, 1), ">=", 26)  # Wednesday
add.constraint(AP_IP_model, c(1, 1, 1, 0, 0, 1, 1), ">=", 25)  # Thursday
add.constraint(AP_IP_model, c(1, 1, 1, 1, 0, 0, 1), ">=", 21)  # Friday
add.constraint(AP_IP_model, c(1, 1, 1, 1, 1, 0, 0), ">=", 19)  # Saturday
```



# Non-Negative constraints:
A1, A2, A3, A4, A5, A6, A7 >=0

# Integral Constraints:
A1, A2, A3, A4, A5, A6, A7 are all integers.

```{r}
# Constrain all the decision variables to integer
set.type(AP_IP_model, 1:7, "integer") 

# Non negative constraints for all the decision variables is already defaultly added in this package.
```

```{r}
# Solve the above integer linear programming problem
solve(AP_IP_model)
```

```{r}
# Get the optimal solution
optimal_solution = get.variables(AP_IP_model)
cat("Optimal solution:", optimal_solution, "\n")
```


# The optimal solution turned out to be 2 workers for Shift 1, 4 workers for Shift 2, 5 workers for Shift 3, 0 workers for Shift 4, 8 workers for Shift 5, 1 worker for Shift 6, and a whopping 13 workers for Shift 7.

```{r}
# Get the total cost
total_cost = get.objective(AP_IP_model)
cat("Total cost:", total_cost, "\n")
```
# Minimum overall expense incurred by the company AP in terms of worker wages is 25675$ subject to the above constraints




