---
title: "rsingav1_4"
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "2023-09-21"
---

# Firstly installing and loading lpSolveAPI library
```{r}
library(lpSolveAPI)
```

#Decision Variables:
L1, M1, S1: Number of large, medium, and small units produced at Plant 1.
L2, M2, S2: Number of large, medium, and small units produced at Plant 2.
L3, M3, S3: Number of large, medium, and small units produced at Plant 3.
There are 9 in number: 
```{r}
# Create a model with 9 decision variables
lprec_weigelt = make.lp(0, 9)
```

# Objective Function: Maximize P = 420*(L1 + L2 + L3) + 360*(M1 + M2 + M3) + 300*(S1 + S2 + S3)

# Define the objective function coefficients (net unit profit). In this case, the net unit profit for each product: Large, Medium and small
```{r}
set.objfn(lprec_weigelt, c(420, 420, 420, 360, 360, 360, 300, 300, 300))
```

# Define the constraint matrix. This includes the production capacities of 3 plants, storage space for each plant, Sales forecast, Same percentage of their excess capacity

# More details on Constraints:

# Capacity Constraints for each plant:

L1 + M1 + S1 <= 750
L2 + M2 + S2 <= 900
L3 + M3 + S3 <= 450

# In-process Storage Space Constraints for each plant:

20*L1 + 15*M1 + 12*S1 <= 13000
20*L2 + 15*M2 + 12*S2 <= 12000
20*L3 + 15*M3 + 12*S3 <= 5000

# Sales Forecast Constraints for each size:

L1 + L2 + L3 <= 900
M1 + M2 + M3 <= 1200
S1 + S2 + S3 <= 750

# Non-negativity Constraints: L1, M1, S1, L2, M2, S2, L3, M3, S3 >= 0

# One last constraint to avoid layoffs is to introduce a condition that the ratio of utilized capacity to total capacity should be the same for all plants. This can be represented as:

(L1 + M1 + S1) / 750 = (L2 + M2 + S2) / 900 = (L3 + M3 + S3) / 450

# This constraint is a bit tricky because it's a ratio, which is nonlinear and cannot be solved by lpSolve. However, we can transform it into a linear form by cross multiplication:

900 * (L1 + M1 + S1) - 750 * (L2 + M2 + S2) =0
450 * (L1 + M1 + S1) - 750 * (L3 + M3 + S3) =0
450 * (L2 + M2 + S2) - 900 * (L3 + M3 + S3) =0

Technically, we can use just two of these three equations as the third equation is just a combination of the first two and the solution would be the same. But I have used all the 3 here just for the sake of completeness.

```{r}
# Adding constraints and defining the constraint directions and right hand side of the constraints as per the question: 
add.constraint(lprec_weigelt, c(1, 0, 0, 1, 0, 0, 1, 0, 0), "<=", 750) # Plant 1 capacity
add.constraint(lprec_weigelt, c(0, 1, 0, 0, 1, 0, 0, 1, 0), "<=", 900) # Plant 2 capacity
add.constraint(lprec_weigelt, c(0, 0, 1, 0, 0, 1, 0, 0, 1), "<=", 450) # Plant 3 capacity
add.constraint(lprec_weigelt, c(20, 0, 0, 15, 0, 0, 12, 0, 0), "<=", 13000) # Plant 1 storage
add.constraint(lprec_weigelt, c(0, 20, 0, 0, 15, 0, 0, 12, 0), "<=", 12000) # Plant 2 storage
add.constraint(lprec_weigelt, c(0, 0, 20, 0, 0, 15, 0, 0, 12), "<=", 5000) # Plant 3 storage
add.constraint(lprec_weigelt, c(1, 1, 1, 0, 0, 0, 0, 0, 0), "<=", 900)# Sales forecast large
add.constraint(lprec_weigelt, c(0, 0, 0, 1, 1, 1, 0, 0, 0), "<=", 1200)# Sales forecast medium
add.constraint(lprec_weigelt, c(0, 0, 0, 0, 0, 0, 1, 1, 1), "<=", 750)# Sales forecast small
add.constraint(lprec_weigelt, c(900, -750, 0, 900, -750, 0, 900, -750, 0), "=", 0) # Same percentage of their excess capacity
add.constraint(lprec_weigelt, c(-450, 0, 750, -450, 0, 750, -450, 0, 750), "=", 0)# Same percentage of their excess capacity
add.constraint(lprec_weigelt, c(0, 450, -900, 0, 450, -900, 0, 450, -900), "=", 0)# Same percentage of their excess capacity
add.constraint(lprec_weigelt, c(1, 0, 0, 0, 0, 0, 0, 0, 0), ">=", 0) # NonNegative constraint
add.constraint(lprec_weigelt, c(0, 1, 0, 0, 0, 0, 0, 0, 0), ">=", 0) # NonNegative constraint
add.constraint(lprec_weigelt, c(0, 0, 1, 0, 0, 0, 0, 0, 0), ">=", 0) # NonNegative constraint
add.constraint(lprec_weigelt, c(0, 0, 0, 1, 0, 0, 0, 0, 0), ">=", 0) # NonNegative constraint
add.constraint(lprec_weigelt, c(0, 0, 0, 0, 1, 0, 0, 0, 0), ">=", 0) # NonNegative constraint
add.constraint(lprec_weigelt, c(0, 0, 0, 0, 0, 1, 0, 0, 0), ">=", 0) # NonNegative constraint
add.constraint(lprec_weigelt, c(0, 0, 0, 0, 0, 0, 1, 0, 0), ">=", 0) # NonNegative constraint
add.constraint(lprec_weigelt, c(0, 0, 0, 0, 0, 0, 0, 1, 0), ">=", 0) # NonNegative constraint
add.constraint(lprec_weigelt, c(0, 0, 0, 0, 0, 0, 0, 0, 1), ">=", 0) # NonNegative constraint

```

```{r}
# Set all variables to be integer as we do not want float values for the unit sizes
for (i in 1:9) {
  set.type(lprec_weigelt, i, "integer")
}
```

# Solve the linear programming problem by maximizing the profit while meeting the constraints

```{r}
# Set the objective function to maximize
lp.control(lprec_weigelt, sense='max')

# Solve the linear programming model
solve(lprec_weigelt)
```

# This will give you the optimal production amounts for each size at each plant to maximize profit, subject to the constraints.
L1 = 530
L2 = 0
L3 = 1
M1 = 160
M2 = 688
M3 = 8
S1 = 0
S2 = 140
S3 = 405
```{r}

# Get the solution
all_unit_sizes_for_all_plants = get.variables(lprec_weigelt)
print(all_unit_sizes_for_all_plants)
```

# Also the optimal profit or maximum profit is: 694680$
```{r}
# Get the optimal profit
maximum_profit = get.objective(lprec_weigelt)
print(maximum_profit) 
```


# Writing the above lp model to a file:
```{r}
write.lp(lprec_weigelt, filename = "weigelt_lp.lp")
```



